---
title: "R Coding Exercise"
output:
  html_document:
    toc: FALSE
---

Import data and begin to clean data using Tidyverse
```{r}
#If needed install package that has our data (dslabs) abd the package that will help us clean the data (tidyverse)
#install.packages(dslabs)
#install.packages(tidyverse)
library(tidyverse)
library(dslabs)

#look at help file for gapminder data
#?gapminder
#get an overview of data structure
str(gapminder)
#get a summary of data
summary(gapminder)
#determine the type of object gapminder is
class(gapminder)

#make a tibble data frame
Data <- gapminder  
as.tibble(Data)
```


Filter all BUT Africa data
```{r}
africadata <- Data %>%
filter(continent == "Africa")
```


Subset and review new data
```{r}
deadbabies <- africadata %>%
select(infant_mortality, life_expectancy)

str(deadbabies)
summary(deadbabies)

popafrica <- africadata %>%
select(population, life_expectancy)

str(popafrica)
summary(popafrica)

```


Plot Data
Population Vs Life Exp
```{r}
#Pop on x-axis
plot1c <- ggplot(data = africadata) + 
  geom_point(mapping = aes(x = population, y =     life_expectancy, color=country)) +
scale_x_continuous(trans = 'log10') 

plot1y <- ggplot(data = africadata) + 
  geom_point(mapping = aes(x = population, y =     life_expectancy, color=year)) +
scale_x_continuous(trans = 'log10') 

plot1c
plot1y
```

Inf Mortality Vs Life Exp
```{r}
plot2y <- ggplot(data = africadata) + 
  geom_point(mapping = aes(x = life_expectancy, y =  infant_mortality, color=year)) +
scale_x_continuous(trans = 'log10') 

plot2c <- ggplot(data = africadata) + 
  geom_point(mapping = aes(x = life_expectancy, y =  infant_mortality, color=country)) +
scale_x_continuous(trans = 'log10') 

plot2c
plot2y
```


Finding all the NAs (ie the missing data)

```{r}
africadata %>%
  summarise(count = sum(is.na(infant_mortality)))

africadata %>%
  summarise(count = sum(is.na(life_expectancy))) 

africadata %>%
  summarise(count = sum(is.na(population))) 
```


Now Filter 2000 data from africadata
```{r}
africa2000data <- africadata %>%
  filter(year == '2000')
```


Plot 2000 data
2000: Population Vs Life Exp
```{r}
plot3c <- ggplot(data = africa2000data) + 
  geom_point(mapping = aes(x = population, y = life_expectancy, color=country)) +
scale_x_continuous(trans = 'log10') 

plot3c
```

2000: Inf Mort Vs Life Exp
```{r}
plot4c <- ggplot(data = africa2000data) + 
  geom_point(mapping = aes(x = life_expectancy, y = infant_mortality, color=country))

plot4c
```

Simple Stats to Evaluate Data
```{r}
fit1 <- lm(infant_mortality ~ life_expectancy, africa2000data)
fit2 <- lm(population ~ life_expectancy, africa2000data)

summary(fit1)
summary(fit2)
```

Visualizing Stats
2000: Population Vs Life Exp
```{r}
ggplot(africa2000data, aes(x = life_expectancy, y = infant_mortality)) + geom_point() +
  stat_smooth(method = "lm", col = "red")
```

2000: Inf Mort Vs Life Exp
```{r}
ggplot(africa2000data, aes(x = life_expectancy, y = population)) + geom_point() +
  stat_smooth(method = "lm", col = "red")
```







