---
title: "Tidy Tuesday Exercise"
output: 
  html_document:
    toc: FALSE
editor: 
  markdown: 
    wrap: 72
---

# Tidy Tuesday Exercise

## Hollywood Age Gaps

### 2/14/2023

Upload public access data for exercise via tidytuesdayR package.

```{r}
#install.packages("tidytuesdayR")
#install.packages('janitor')

library(tidytuesdayR)
library(tidyverse)
library(here)
library(janitor)

tuesdata <- tidytuesdayR::tt_load(2023, week = 7)
age_gaps <- tuesdata$age_gaps

```

## Data Dictionary

### variable; class; description

movie_name; character; Name of the film;

release_year; integer; Release year

director; character; Director of the film

age_difference; integer; Age difference between the characters in whole
years

couple_number; integer; An identifier for the couple in case multiple
couples are listed for this film

actor_1\_name; character; The name of the older actor in this couple

actor_2\_name; character; The name of the younger actor in this couple

character_1\_gender; character; The gender of the older character, as
identified by the person who submitted the data for this couple

character_2\_gender; character; The gender of the younger character, as
identified by the person who submitted the data for this couple

actor_1\_birthdate; date; The birthdate of the older member of the
couple

actor_2\_birthdate; date; The birthdate of the younger member of the
couple

actor_1\_age; integer; The age of the older actor when the film was
released

actor_2\_age; integer; The age of the younger actor when the film was
released

```{r}
tibble(age_gaps)

glimpse(age_gaps)
```

# Cleaning data

```{r}
adata <- age_gaps %>%
  ##Pick out data relevant to me
  ## filter out main couples (i.e. couple 1)
  filter(couple_number == 1) %>%
  ## select variables of interest
  select( release_year  | age_difference | movie_name | character_1_gender |  actor_1_age | actor_1_name | character_2_gender | actor_2_age | actor_2_name )

##quick visualization of data to look for trends
plot(adata)

#quick view of data with scatterplots show that actor 1 age and age difference seem related
#but i think this data would be easier to read if we changed actor gender to just male vs female
#we'll flip this to female lead roles and male lead roles

leadFdata <- adata %>%
  filter(character_1_gender == 'woman') %>%
  rename("female_leads" = "character_1_gender") %>%
  rename("lead_female_age" = "actor_1_age") %>%
  rename("lead_name" = "actor_1_name") %>%
  rename("supporting_name" = "actor_2_name") %>%
  rename("supporting_actor_age" = "actor_2_age") %>%
  rename("supporting_gender" = "character_2_gender")

plot(leadFdata)

#looks promising to pursue: age difference and lead female age, lead age and supporting actor age, also personally interested in supporting actress age and release year

leadMdata <- adata %>%
  filter(character_1_gender == 'man') %>%
  rename("male_leads" = "character_1_gender") %>%
  rename("lead_male_age" = "actor_1_age")  %>%
  rename("lead_name" = "actor_1_name") %>%
  rename("supporting_name" = "actor_2_name") %>%
  rename("supporting_actor_age" = "actor_2_age") %>%
  rename("supporting_gender" = "character_2_gender")

plot(leadMdata)

#firstly wow! way more data. Second same trends seem to apply. Would also like to compare the gender of the supporting actor and a same-sex supporting role.
#onto visualization!

```
# Data Vizualization

## Actor and Release Data
### (because inquiring minds want to know...)


```{r}
library(ggthemes)

ARplot <- ggplot() +
  geom_point(data = leadMdata, aes(x = release_year, y = lead_male_age), color = 'dodgerblue4', size=1.5, shape = 15) +
  geom_point(data = leadFdata, aes(x = release_year, y = lead_female_age), color = 'deeppink4', size=1.5, shape = 15) +
    geom_point(data = leadMdata, aes(x = release_year, y = supporting_actor_age), color = 'deepskyblue2', size=1.5, shape = 18) +
  geom_point(data = leadFdata, aes(x = release_year, y = supporting_actor_age), color = 'deeppink1', size=1.5, shape = 18) +
  ggtitle("Actor Age in Relation to the Movie Release Data", subtitle = "Evaluated by Genders") +
labs(x = "Release Year", y = "Actor Ages") +
  annotate(geom="text", x=1950, y=75, label="Lead Male Actors", colour="dodgerblue4", size=4, family="sans", fontface="bold", angle=0) +
  annotate(geom="text", x=1950, y=80, label="Lead Female Actors", colour="deeppink4", size=4, family="sans", fontface="bold", angle=0) +
  annotate(geom="text", x=1950, y=65, label="Actors that Support Male Leads", colour="deepskyblue2", size=3, family="sans", fontface="bold", angle=0) +
  annotate(geom="text", x=1950, y=70, label="Actors that Support Female Leads", colour="deeppink1", size=3, family="sans", fontface="bold", angle=0)

ARplot

```
Lets look a little closer at lead actors and age gaps

```{r}
ARAplot <- ggplot() +
  geom_point(data = leadMdata, aes(x = release_year, y = lead_male_age, size = age_difference), fill = 'dodgerblue4', shape = 21, colour = "black") +
  geom_point(data = leadFdata, aes(x = release_year, y = lead_female_age, size = age_difference), fill = 'deeppink4', shape = 21, colour = "black") +

  ggtitle("Lead Actor Age in Relation to the Movie Release Data", subtitle = "Factoring in Age Gaps between Lead and Supporting Actors") +
  labs(x = "Release Year", y = "Actor Ages", color= "Age Gap") +
  annotate(geom="text", x=1950, y=75, label="Lead Male Actors", colour="dodgerblue4", size=4, family="sans", fontface="bold", angle=0) +
  annotate(geom="text", x=1950, y=80, label="Lead Female Actors", colour="deeppink4", size=4, family="sans", fontface="bold", angle=0) 
  


ARAplot

```


## Relationship between Movie release data, Age difference, and supporting Actor Age 
###Supporting actor age

```{r}
ASAplot <- ggplot() +
    geom_point(data = leadMdata, aes(x = release_year, y = supporting_actor_age, size = age_difference), fill = 'deepskyblue2', shape = 21, colour = "black") +
  geom_point(data = leadFdata, aes(x = release_year, y = supporting_actor_age, size = age_difference), fill = 'deeppink1', shape = 21, colour = "black") +
  ggtitle("Supporting Actor Age in Relation to the Movie Release Data", subtitle = "Factoring in Age Gaps between Lead and Supporting Actors") +
labs(x = "Release Year", y = "Actor Ages") +
   annotate(geom="text", x=1965, y=60, label="Actors that Support Male Leads", colour="deepskyblue2", size=4, family="sans", fontface="bold", angle=0) +
  annotate(geom="text", x=1965, y=65, label="Actors that Support Female Leads", colour="deeppink1", size=4, family="sans", fontface="bold", angle=0)

ASAplot
```

## Lead Age vs Supporting actor age
### Opposite vs Same-sex
#### lets add some trends

Males:
```{r}
SSplot <- ggplot() +
  geom_point(data = leadMdata, aes(x = age_difference, y = lead_male_age), color = 'dodgerblue4', size=2) +
  geom_point(data = leadMdata, aes(x = age_difference, y = supporting_actor_age, color = supporting_gender), size=2 , shape = 21) +
  ggtitle("Male Actor Age in Relation to the Age Gaps", subtitle = "Evaluated by Supporting Actor Genders") +
  labs(x = "Age Difference Between Supporitng and Lead Actors", y = "Actor Ages") +
  annotate(geom="text", x=7, y=80, label="Lead Male Actors", colour="dodgerblue4", size=4, family="sans", fontface="bold", angle=0)  + 
  annotate(geom="text", x=7, y=75, label="Supporting Actors", color="cyan4", size=4, family="sans", fontface="bold", angle=0) 


SSplot 
```
Females:
```{r}
SSFplot <- ggplot() +
  geom_point(data = leadFdata, aes(x = age_difference, y = lead_female_age), color = 'deeppink4', size=2) +
  geom_point(data = leadFdata, aes(x = age_difference, y = supporting_actor_age, color = supporting_gender), size=2 , shape = 21) +
  ggtitle("Female Actor Age in Relation to the Age Gaps", subtitle = "Evaluated by Supporting Actor Genders") +
  labs(x = "Age Difference Between Supporitng and Lead Actors", y = "Actor Ages") +
  annotate(geom="text", x=10, y=75, label="Lead Female Actors", colour="deeppink4", size=4, family="sans", fontface="bold", angle=0)  + 
  annotate(geom="text", x=10, y=70, label="Supporting Actors", color="orangered2", size=4, family="sans", fontface="bold", angle=0) 


SSFplot 
```
Wow this is pretty striking! The raw cleaned data really shows a disparity between male leads and female leads and the age gap between their supporting actors!

We could play with this data forever! I will stop here but if you have a recommendation please post it on my "issues" page of my github!