# Improving Models

### Load packakes and data

```{r}
library(here)
library(tidyverse)
library(ggplot2)
library(tidymodels)
library(dplyr)

data <- readRDS(here('fluanalysis/data/SypAct_clean.rds')) #upload cleaned data
tibble(data) #overview of data
```

### Split Data

```{r}
set.seed(123)
# Fix the random numbers by setting the seed 
# This enables the analysis to be reproducible when random numbers are used 

data_split <- initial_split(data, prop = 3/4)

# Create data frames for the two sets:
train_data <- training(data_split)
test_data  <- testing(data_split)

```

### Make a recipe and fit a model

```{r}
# Create recipe using Nausea as categorical variable
data_recipe <- recipe(Nausea ~ ., data = train_data) 

# Logistic model recipe
recipe_mod <- logistic_reg() %>% set_engine("glm")

# Model workflow to pair model and recipe 
mod_flow <- workflow() %>% 
  add_model(recipe_mod) %>% 
  add_recipe(data_recipe)

mod_flow
```

### Train the model from the resulting predictors

```{r}
data_fit <- 
  mod_flow %>% 
  fit(data = train_data)

data_fit %>% 
  extract_fit_parsnip() %>% 
  tidy()

```

### Use a trained workflow to Predict

```{r}
predict(data_fit, test_data)

data_aug <- 
  augment(data_fit, test_data)

# The data look like: 
data_aug %>%
  select(Nausea, .pred_No, .pred_Yes)

```

### ROC Curve

```{r}
data_aug %>% 
  roc_curve(truth = Nausea, .pred_No) %>% 
  autoplot()

data_aug %>% 
  roc_auc(truth = Nausea, .pred_No)
# > 0.7; the model might be useful. 
```

### Alternative Model: Main predictor to the Categorical Outcome

```{r}

# Create new recipe
data_recipe2 <- recipe(Nausea ~ Myalgia, data = train_data) 

# Model workflow to pair model and recipe 
mod_flow2 <- workflow() %>% 
  add_model(recipe_mod) %>% 
  add_recipe(data_recipe2)

#Train data
data_fit2 <- 
  mod_flow2 %>% 
  fit(data = train_data)

data_aug2 <- 
  augment(data_fit2, test_data)

data_aug2 %>% 
  roc_auc(truth = Nausea, .pred_No)
# < 0.7; the model not very useful. 

```


### The following added by SETH LATTNER

#### Fitting continuous variables

```{r}
# Create recipe using Nausea as categorical variable
data_recipe3 <- recipe(BodyTemp ~ ., data = train_data) 

# Logistic model recipe
recipe_mod3 <- linear_reg() %>% set_engine("lm")

# Model workflow to pair model and recipe 
mod_flow3 <- workflow() %>% 
  add_model(recipe_mod3) %>% 
  add_recipe(data_recipe3)

mod_flow3
```

### Train the model from the resulting predictors

```{r}
data_fit3 <- 
  mod_flow3 %>% 
  fit(data = train_data)

data_fit3 %>% 
  extract_fit_parsnip() %>% 
  tidy()

```

### Use a trained workflow to Predict

```{r}
predict(data_fit3, test_data)

data_aug3 <- 
  augment(data_fit3, test_data)

# The data look like: 
data_aug3 %>%
  select(BodyTemp, .pred)

```
```{r}
#calculate RMSE
yardstick::rmse(data_aug3, BodyTemp, .pred)
```


### Alternative Model: Main predictor to the Categorical Outcome

```{r}

# Create new recipe
data_recipe4 <- recipe(BodyTemp ~ RunnyNose, data = train_data) 

# Model workflow to pair model and recipe 
mod_flow4 <- workflow() %>% 
  add_model(recipe_mod3) %>% 
  add_recipe(data_recipe4)

#Train data
data_fit4 <- 
  mod_flow4 %>% 
  fit(data = train_data)

data_fit4 %>% 
  extract_fit_parsnip() %>% 
  tidy()
```

```{r}
#predict from test data
predict(data_fit4, test_data)

#augment test data
data_aug4 <- 
  augment(data_fit4, test_data)

#the data look like:
data_aug4 %>%
  select(BodyTemp, .pred)

```
```{r}
#calculate RMSE
yardstick::rmse(data_aug4, BodyTemp, .pred)
```
The RMSE of the univariate model 4 (1.12) was lower than that of the global model 3 (1.15), showing that it is a better fit to the data.
