# **Flu Analysis: Wrangling**

#### Load the data and packages:

```{r}
library(tidyverse)
library(here)

raw_flu <- readRDS(here('fluanalysis/data/SympAct_Any_Pos.Rda'))
tibble(raw_flu) # to get a look at the data
```

#### Remove unwanted data columns:

Remove all variables that have Score or Total or FluA or FluB or Dxname or Activity in their name. Also remove the variable Unique.Visit. Remove any NA observations.

Now data set contains 32 variables coding for presence or absence of some symptom. Only one, temperature, is continuous.

```{r}
flu_clean <- raw_flu %>% #sort through the data and removing all variables (columns) that include the words: Score or Total or FluA or FluB or Dxname or Activity
  select(-contains(c("Score", "Total", "FluA", "FluB", "Dxname", "Activity"))) %>% #now will remove all columns that include Unique.Visit
  select(-contains(c("Unique.Visit"))) %>%
  drop_na #Will drop all nas now


```

#### Save file into project:

```{r}
saveRDS(flu_clean, file= here("fluanalysis", "data", "SypAct_clean.rds")) #will save as a data.frame in the RDS file

```
